---
title: "Google Analytics and Machine Learning 2nd Part"
author: "Marco Russo"
date: "12 de mayo de 2019"
output: html_document
---

```{r message=FALSE}
# Instalamos las librerias
library("forecast")
library("reshape2")
library("stats")
library("ggplot2")
library("ggthemes")
library("ggrepel")
library(dplyr)
library(tidyr)


# Use google_analytics_3 old version and 
library(googleAnalyticsR)
# library(RGoogleAnalytics) Alternative with secretID and Token
library(ggplot2) # For some initial plots.
library(forecast) # for the time series prediction.
library(tidyverse)
library(bupaR)
```


```{r,eval=TRUE,echo=TRUE,warning=FALSE, message=FALSE}
## Autorizamos a través del token con nuestro account Google
ga_auth()



```{r message=FALSE}
## Veamos el listado de los account de GA y la guardamos en una nueva variable
account_list <- ga_account_list()
# comprobamos la estructura
str(account_list)

## visualizamos la columna Vista, viewId
head(account_list$viewId, n = 10)

## se recoge la vista del índice 4 y se muestra
ga_id <- account_list[4,'viewId']
ga_id

```



## Create dataframe with main metrics and dimensions


```{r include=FALSE}
# Creamos dataframe con diferentes variables
web_data_metrics <- google_analytics_3(id = ga_id,
                                       start="2018-01-01", end="2018-06-30",
                                       metrics = c("sessions","pageviews","entrances","bounces"),
                                       dimensions = c("date","channelGrouping","deviceCategory"),
                                       max = 5000,
                                       samplingLevel = 'WALK')

```




```{r}
## agrupamos solo las dimensiones fecha, canales y dispositivos en desktop
pivoted <- web_data_metrics %>% 
  filter(deviceCategory == "desktop") %>% 
  select(date, channelGrouping, sessions) %>%
  spread(channelGrouping, sessions)

## pivotamos realizando el cambio de NA's a ceros
pivoted[is.na(pivoted)] <- 0

head(pivoted)
```


## A seasonal time-series


```{r}
## creamos una serie temporal con frecuencia semanal
web_data_ts <- ts(pivoted[-1], frequency = 7)

## Creamos un plot 
plot(web_data_ts, axes = FALSE)

```


### Time-series appplications
Decomposition
With seasonal data, you can extract the seasonality (weekly in this example) so you can more easily see the overall trend of data:
```{r}
decomp <- decompose(web_data_ts[, "Organic Search"])
plot(decomp)
```

- Observed - the original data
- Trend - the data minus the seasonal
- Seasonal - data that fits the specified seasonality (7 days)
- Random - everything else left only - useful for anomoloy detection.



```{r}
library(forecast)
## performs decomposition and smoothing
fit <- ets(web_data_ts[, "Organic Search"])
## makes the forecast
fc <- forecast(fit)
plot(fc)

```



## From Holtwinters
```{r}

fit2 <- HoltWinters(web_data_ts[, "Organic Search"])
## makes the forecast
fc2 <- forecast(fit2, h = 25)
plot(fc2)

```





# Forecasting 26 días

```{r message=FALSE}
# 
gadata_5 <- google_analytics_3(id = ga_id, 
                           start="2018-06-01", end="2018-06-30",
                           metrics = "sessions", 
                           dimensions = "date",
                           max = 5000)


timeseries <- ts(gadata_5$sessions, frequency=7)
components <- decompose(timeseries)
plot(components)

# note the way we add a column to a data.frame
gadata_5$adjusted <- gadata_5$sessions - components$seasonal

theme(axis.text.x = element_text(angle = 90, hjust = 1))

forecastmodel <- HoltWinters(timeseries)
plot(forecastmodel)

forecast <- forecast:::forecast.HoltWinters(forecastmodel, h=26) # 26 days in future
plot(forecast, xlim=c(0,12))

forecastdf <- as.data.frame(forecast)
totalrows <- nrow(gadata_5) + nrow(forecastdf)
forecastdata <- data.frame(day=c(1:totalrows),
actual=c(gadata_5$sessions,rep(NA,nrow(forecastdf))),
forecast=c(rep(NA,nrow(gadata_5)-1),tail(gadata_5$sessions,1),forecastdf$"Point Forecast"),
forecastupper=c(rep(NA,nrow(gadata_5)-1),tail(gadata_5$sessions,1),forecastdf$"Hi 80"),
forecastlower=c(rep(NA,nrow(gadata_5)-1),tail(gadata_5$sessions,1),forecastdf$"Lo 80")
)

ggplot(forecastdata, aes(x=day)) +
geom_line(aes(y=actual),color="black") +
geom_line(aes(y=forecast),color="blue") +
geom_ribbon(aes(ymin=forecastlower,ymax=forecastupper), alpha=0.4, fill="green") +
xlim(c(0,90)) +
xlab("Day") +
ylab("Sessions")
```

# Forecasting 90 días

```{r message=FALSE}
library("forecast")
library("reshape2")
library("stats")

gadata_6 <- google_analytics_3(id = ga_id, 
                           start="2019-01-01", end="2019-01-30",
                           metrics = "sessions", 
                           dimensions = "date",
                           max = 5000)


timeseries <- ts(gadata_6$sessions, frequency=7)
components <- decompose(timeseries)
plot(components)

# note the way we add a column to a data.frame
gadata_6$adjusted <- gadata_6$sessions - components$seasonal

theme(axis.text.x = element_text(angle = 90, hjust = 1))

forecastmodel <- HoltWinters(timeseries)
plot(forecastmodel)

forecast <- forecast:::forecast.HoltWinters(forecastmodel, h=30) # 26 days in future
plot(forecast, xlim=c(0,13))

forecastdf <- as.data.frame(forecast)
totalrows <- nrow(gadata_6) + nrow(forecastdf)
forecastdata <- data.frame(day=c(1:totalrows),
actual=c(gadata_6$sessions,rep(NA,nrow(forecastdf))),
forecast=c(rep(NA,nrow(gadata_6)-1),tail(gadata_6$sessions,1),forecastdf$"Point Forecast"),
forecastupper=c(rep(NA,nrow(gadata_6)-1),tail(gadata_6$sessions,1),forecastdf$"Hi 50"),
forecastlower=c(rep(NA,nrow(gadata_6)-1),tail(gadata_6$sessions,1),forecastdf$"Lo 50")
)

ggplot(forecastdata, aes(x=day)) +
geom_line(aes(y=actual),color="black") +
geom_line(aes(y=forecast),color="blue") +
geom_ribbon(aes(ymin=forecastlower,ymax=forecastupper), alpha=0.4, fill="green") +
xlim(c(0,50)) +
xlab("Day") +
ylab("Sessions")
```


### Converting to xts for some packages
We can only work with numeric data, but we would also like to keep the date information, mainly for labelling purposes. xts lets you create the time-series whilst specifying the date labels like so:

```{r}
library(xts)

## create a time-series zoo object
web_data_xts <- xts(pivoted[-1], order.by = as.Date(pivoted$date), frequency = 7)

```

web_data_xts will look similar to when it was a data.frame, but its now in the right class to do some nice related functions.

### CausalImpact
A good example of a library that needs xts is CausalImpact, which is a fantastic library from Google well suited for digital marketing.

CausalImpact gives you an estimate on how much effect an event at a certain point of time had on your metrics, in absolute and relative terms. In this case, an event could be a TV campaign starting or changing your Title tags. It also lets you add control segments, so you can adjust for known effects.

As an example, we assume the observed peak in Social traffic is due to the start of some campaign in May 15th 2016 (in reality you should not cherry pick dates like this!) and we woud like to observe its effect on Video sessions. We also add Direct sessions as a control to help account for general website trends.


```{r}
# install.packages(CausalImpact)
library(CausalImpact)
library(bupaR)
pre.period <- as.Date(c("2018-03-13","2018-04-14"))
post.period <- as.Date(c("2018-04-15","2018-06-30"))

## data in order of response, predictor1, predictor2, etc.
model_data <- web_data_xts[,c("Direct","Social","Organic Search","Paid Search","Referral")]


impact <- CausalImpact(model_data,  pre.period, post.period)
plot(impact)


```

- original - the original response data, with a forecast (blue) of where it should have been in post event period
- pointwise - the difference between the forecast and reality per day
- cumulative - the cumuative effect on sessions in the post event period



## Correlation


```{r}
## see correlation between all metrics
kable(cor(web_data_metrics[,c("sessions","pageviews","entrances","bounces")]))
```


```{r}
## see correlation between all metrics
pairs(web_data_metrics[,c("sessions","pageviews","entrances","bounces")])
```


```{r}
## see correlation between all metrics
cor(web_data_metrics[,c("sessions","pageviews","entrances","bounces")])
```


## How do web channels correlate?
One useful piece of analysis is seeing how web channels possibly interact

```{r}
## Use tidyverse to pivot the data
library(dplyr)
library(tidyr)
library(knitr)

# Create un nuevo dataset
gadata_8 <- google_analytics_3(id = ga_id, 
                           start="2018-01-01", end="2018-06-30",
                           metrics = "sessions", 
                           dimensions = c("date","deviceCategory","channelGrouping"),
                           max = 5000)



## Get only desktop rows, and the date, channelGrouping and sessions columns
pivoted <- web_data_metrics %>% 
  filter(deviceCategory == "desktop") %>% 
  dplyr::select(date, channelGrouping, sessions) %>%
  spread(channelGrouping, sessions)

## Get rid of any NA's and replace with 0
pivoted[is.na(pivoted)] <- 0

kable(head(pivoted))


```


```{r}
## can't include the date as its not numeric, so remove
cor_data <- pivoted[, -1]
## not including first column, so -1 subset
cor_table <- round(cor(cor_data),2)
kable(cor_table)

```

```{r}
## can't include the date as its not numeric, so remove
pairs(cor_data)

```

## Analysis

Now, when we compare channels, we see much looser correlations for this dataset, which makes sense, right? Correlations under 0.3 are, as a rule-of-thumb, not worth considering, so the standouts look to be Social vs. Video* and Paid** vs. Organic Search.

Plotting those channels, we can examine the trends to see the shape of the data

Correlation has help us zero in on possibly interesting relationships


```{r}
library(ggplot2)
gg <- ggplot(data = pivoted) + 
      theme_minimal() + 
      ggtitle("Pago (blue) vs Orgánico (green)")
gg <- gg + 
      geom_line(aes(x = as.Date(date), y = `Paid Search`), col = "blue")

gg + geom_line(aes(x = as.Date(date), y = `Organic Search`), col = "green")

``` 

```{r}
library(ggplot2)
gg <- ggplot(data = pivoted) + 
              theme_minimal() + 
              ggtitle("Social (red) vs Referral (orange)")
gg <- gg + 
      geom_line(aes(x = as.Date(date), y = Social), col = "red")
gg + geom_line(aes(x = as.Date(date), y = Referral), col = "orange")

```


```{r}
library(ggplot2)
gg <- ggplot(data = pivoted) + 
              theme_minimal() + 
              ggtitle("Orgánico(verde) vs Referral (orange)")
gg <- gg + 
      geom_line(aes(x = as.Date(date), y = `Organic Search`), col = "green")
gg + geom_line(aes(x = as.Date(date), y = Referral), col = "orange")

```


```{r}
library(ggplot2)
gg <- ggplot(data = pivoted) + 
              theme_minimal() + 
              ggtitle("Directo (gris) vs Paid (blue)")
gg <- gg + 
      geom_line(aes(x = as.Date(date), y = `Direct`), col = "grey")
gg + geom_line(aes(x = as.Date(date), y = `Paid Search`), col = "blue")

```


```{r}
library(ggplot2)
gg <- ggplot(data = pivoted) + 
              theme_minimal() + 
              ggtitle("Directo (gris) vs Display (blue)")
gg <- gg + 
      geom_line(aes(x = as.Date(date), y = `Direct`), col = "grey")
gg + geom_line(aes(x = as.Date(date), y = `Display`), col = "blue")

```





```{r}
ccf(pivoted$Social, pivoted$Referral)

``` 

```{r}
library(dplyr) # data transformation

head(web_data_metrics)

cats <- web_data_metrics %>% dplyr::select(deviceCategory, channelGrouping, sessions)


## Get only desktop rows, and the date, channelGrouping and sessions columns
pivoted <- gadata_8 %>% 
  filter(deviceCategory == "desktop") %>% 
  dplyr::select(date, channelGrouping, sessions) %>%
  spread(channelGrouping, sessions)

## Get rid of any NA's and replace with 0
pivoted[is.na(pivoted)] <- 0

kable(head(pivoted))



kable(head(cats), row.names = FALSE)

```


```{r}
library(dplyr) # data transformation
cats <- web_data_metrics %>% dplyr::select(deviceCategory, channelGrouping, sessions)

kable(head(cats), row.names = FALSE)

```

```{r}
library(rpart) # creates decision trees

tree <- rpart(deviceCategory ~ ., cats)
plot(tree)
text(tree)
```




```{r}
library(rpart.plot)  # pretty trees
rpart.plot(tree, type=2) 

```











